Bow shocks, bow waves, and dust waves. I. Strong coupling limit

### 2019-04-07 papers/Papers1234/bs-bw-dw-01.tex
Finish first complete draft of Paper I
Submit BS-BW-DW-01 paper
Finish intro for Paper III
Submit revised version of Paper I to MNRAS
Expanded captions for Figs 5, 6, 7.  Change "bow" to "bow shell" in
many places.

\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle
\begin{abstract}
  Dust waves and bow waves result from the action of a star's
  radiation pressure on a stream of dusty plasma that flows past it.
  They are an alternative mechanism to hydrodynamic bow shocks for
  explaining the curved arcs of infrared emission seen around some
  stars.  When gas and grains are perfectly coupled, for a broad class
  of stellar parameters, wind-supported bow shocks predominate when
  the ambient density is below \SI{100}{cm^{-3}}.  At higher densities
  radiation-supported bows can form, tending to be optically thin bow
  waves around B~stars, or optically thick bow shocks around early
  O~stars.  For OB stars with particularly weak stellar winds,
  radiation-supported bows become more prevalent.
\end{abstract}

@@

\input{sec-dust-bow-intro}
\input{sec-strong-coupling}
\bibliographystyle{mnras}
\bibliography{bowshocks-biblio}

>>>>

\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle
\begin{abstract}
  Dust waves and bow waves result from the action of a star's
  radiation pressure on a stream of dusty plasma that flows past it.
  They are an alternative mechanism to hydrodynamic bow shocks for
  explaining the curved arcs of infrared emission seen around some
  stars.  When gas and grains are perfectly coupled, for a broad class
  of stellar parameters, wind-supported bow shocks predominate when
  the ambient density is below \SI{100}{cm^{-3}}.  At higher densities
  radiation-supported bow shells can form, tending to be optically
  thin bow waves around B~stars, or optically thick bow shocks around
  early O~stars.  For OB stars with particularly weak stellar winds,
  radiation-supported bow shells become more prevalent.
\end{abstract}

@@

\input{sec-dust-bow-intro}
\input{sec-strong-coupling}

\section*{Acknowledgements}
We are grateful for financial support provided by Dirección General de
Asuntos del Personal Académico, Universidad Nacional Autónoma de
México, through grants Programa de Apoyo a Proyectos de Investigación
e Inovación Tecnológica IN112816 and IN107019.  We thank the referee
for a helpful report.

\bibliographystyle{mnras}
\bibliography{bowshocks-biblio}

### 2019-04-07 papers/Papers1234/sec-dust-bow-intro.tex
Jane's Paper I corrections, up to end of sec 2.3
Submit revised version of Paper I to MNRAS
Expanded captions for Figs 5, 6, 7.  Change "bow" to "bow shell" in
many places.

\begin{figure*}
  \centering
  \includegraphics[width=0.8\linewidth]{figs/bows-and-waves}
  \caption{Regimes in the supersonic interaction of a luminous star
    with its environment.  When radiation effects are unimportant, we
    have the standard (magneto-)hydrodynamical wind-supported bow
    shock (upper left), where the ram pressure of the stellar wind
    balances the ram pressure of the oncoming stream.  As the optical
    depth of the shocked shell increases, the stellar radiation
    momentum adds to the wind ram pressure to help support the bow.
    If the shell is completely opaque to stellar radiation, we have a
    radiation-supported bow shock (lower right), where it is the
    stellar radiation pressure that balances the ram pressure of the
    external stream.  For intermediate optical depths, we have two
    cases depending on the strength of coupling between grains and
    gas.  If the coupling is strong, then we have a
    radiation-supported bow wave (upper right), where the plasma
    stream as a whole is gradually radiatively decelerated as it
    approaches the star.  If the coupling is weak, then the radiation
    momentum is felt only by the dust, which decouples from the gas to
    form a dust wave, with the gas stream continuing inward to form a
    wind-supported bow shock closer to the star.  Note that the
    gas--grain coupling is both direct (via collisions) and indirect
    (via the magnetic field) (see Paper~II).}
  \label{fig:3-types-bow}
\end{figure*}

@@

This is the first in a series of papers where we develop simple
physical models to show in detail when and how these different
interaction regimes apply when varying the parameters of the star, the
dust grains, and the ambient stream.  We concentrate primarily on the
case of luminous early type stars, where dust is present only in the
ambient stream, and not in the stellar wind.  In this first paper, we
consider the case where the grains are perfectly coupled to the gas
via collisions.  The following two papers consider the decoupling of
grains and gas in a sufficiently strong radiation field
\citep[Paper~II]{Henney:2019b}, and how observations can distinguish
between different classes of bow \citep[Paper~III]{Henney:2019c}.
%
The paper is organized as follows.
%
In \S~\ref{sec:strong-gas-grain} we propose a simple model for stellar
bows and investigate the relative importance of wind and radiation in
providing internal support for the bow shell as a function of the
density and velocity of the ambient stream, and for different types of
star.  In \S~\ref{sec:phys-state-shock} we calculate the physical
state of the bow shell, considering under what circumstances it can
trap within itself the star's ionization front and how efficient
radiative cooling will be.
%
In \S~\ref{sec:discussion} we briefly discuss the application
of our models to observed bows.
%
In \S~\ref{sec:summary} we summarise the conclusions of our study. 
%

>>>>

\begin{figure*}
  \centering
  \includegraphics[width=0.8\linewidth]{figs/bows-and-waves}
  \caption{Regimes in the supersonic interaction of a luminous star
    with its environment.  When radiation effects are unimportant, we
    have the standard (magneto-)hydrodynamical wind-supported bow
    shock (upper left), where the ram pressure of the stellar wind
    balances the ram pressure of the oncoming stream.  As the optical
    depth of the shocked shell increases, the stellar radiation
    momentum adds to the wind ram pressure to help support the bow
    shell.  If the shell is completely opaque to stellar radiation, we
    have a radiation-supported bow shock (lower right), where it is
    the stellar radiation pressure that balances the ram pressure of
    the external stream.  For intermediate optical depths, we have two
    cases depending on the strength of coupling between grains and
    gas.  If the coupling is strong, then we have a
    radiation-supported bow wave (upper right), where the plasma
    stream as a whole is gradually radiatively decelerated as it
    approaches the star.  If the coupling is weak, then the radiation
    momentum is felt only by the dust, which decouples from the gas to
    form a dust wave, with the gas stream continuing inward to form a
    wind-supported bow shock closer to the star.  Note that the
    gas--grain coupling is both direct (via collisions) and indirect
    (via the magnetic field) (see Paper~II).}
  \label{fig:3-types-bow}
\end{figure*}

@@

This is the first in a series of papers where we develop simple
physical models to show in detail when and how these different
interaction regimes apply when varying the parameters of the star, the
dust grains, and the ambient stream.  We concentrate primarily on the
case of luminous early type stars, where dust is present only in the
ambient stream, and not in the stellar wind.  In this first paper, we
consider the case where the grains are perfectly coupled to the gas
via collisions.  The following two papers consider the decoupling of
grains and gas in a sufficiently strong radiation field
\citep[Paper~II]{Henney:2019b}, and how observations can distinguish
between different classes of bow shell
\citep[Paper~III]{Henney:2019c}.
%
The paper is organized as follows.
%
In \S~\ref{sec:strong-gas-grain} we propose a simple model for stellar
bow shells and investigate the relative importance of wind and radiation in
providing internal support for the bow shell as a function of the
density and velocity of the ambient stream, and for different types of
star.  In \S~\ref{sec:phys-state-shock} we calculate the physical
state of the bow shell, considering under what circumstances it can
trap within itself the star's ionization front and how efficient
radiative cooling will be.
%
In \S~\ref{sec:discussion} we briefly discuss the application
of our models to observed bow shells.
%
In \S~\ref{sec:summary} we summarise our findings. 
%
